---
title: "BOAT2-Cluster-and-Factor"
output: html_document
---

# Data Analysis and Clustering of Building Owners' Decision-Making Profiles

## Load Required Libraries
```{r}
# Core data manipulation and visualization
library(tidyverse)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(scales)

# Clustering and dimension reduction
library(cluster)
library(clustMixType)
library(factoextra)
library(NbClust)
library(umap)

# Visualization enhancements
library(dendextend)
library(viridis)
library(RColorBrewer)
library(ggrepel)
```

## Data Preparation and Initial Clustering Analysis
This section focuses on preparing the data and performing hierarchical clustering analysis to identify distinct groups of building owners based on their decision-making profiles.

```{r}
# Read the data
boat_data <- read.csv("BOAT2_Data.csv")

# Create a unique ID for each observation
boat_data$ID <- 1:nrow(boat_data)

# Extract and process the decision-making profile characteristics
dmpc_data <- boat_data %>%
  # Calculate mean values for multi-item constructs
  mutate(
    Distribution_Centralization = (Distribution_Centralization1 + Distribution_Centralization2) / 2,
    Distribution_Formalization = (Distribution_Formalization1 + Distribution_Formalization2) / 2,
    Style_Participation = (Style_Participation1 + Style_Participation2) / 2,
    Style_Organicity = (Style_Organicity1 + Style_Organicity2) / 2,
    Style_Coercion = (Style_Coercion1 + Style_Coercion2) / 2
  ) %>%
  # Select the variables for clustering
  select(
    ID, 
    Distribution_Centralization, Distribution_Formalization,
    Style_Technocracy, Style_Participation, Style_Organicity, Style_Coercion,
    Culture_Command, Culture_Symbolic, Culture_Rationale, Culture_Generative, Culture_Transactive,
    Flexibility_openness, Flexibility_Recursiveness,
    Risk,
    Environment_Growth, Environment_Hostile, Environment_Stable,
    PDM_Type
  )

# Handle missing values
dmpc_data <- dmpc_data %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), mean(., na.rm = TRUE), .)))

# Extract PDM_Type for later use
pdm_types <- dmpc_data$PDM_Type
ids <- dmpc_data$ID

# Prepare data for clustering (only numeric variables)
dmpc_matrix <- dmpc_data %>%
  select(-ID, -PDM_Type) %>%
  as.matrix()

# Scale the data
dmpc_scaled <- scale(dmpc_matrix)
rownames(dmpc_scaled) <- ids

# Compute distance matrix
dist_matrix <- dist(dmpc_scaled, method = "euclidean")

# Perform hierarchical clustering using Ward's method
# Ward's method tends to create more balanced, interpretable clusters
hc_ward <- hclust(dist_matrix, method = "ward.D2")

# Create a dendrogram object
dend <- as.dendrogram(hc_ward)

# Define color palette for PDM types
pdm_colors <- brewer.pal(length(unique(pdm_types)), "Set1")
names(pdm_colors) <- unique(pdm_types)

# Color the labels based on PDM type
pdm_colors_mapped <- pdm_colors[pdm_types]
labels_colors(dend) <- pdm_colors_mapped[order.dendrogram(dend)]

# Set up the plot with larger dimensions and margins
pdf("BOAT2_Dendrogram_PDM.pdf", width = 12, height = 8)
par(mar = c(8, 4, 4, 10))  # Bottom, left, top, right margins

# Plot the dendrogram
plot(dend,
     main = "Hierarchical Clustering of Building Owners by Decision-Making Profiles",
     sub = "Colored by Project Delivery Method (PDM)",
     xlab = "",
     ylab = "Height (Dissimilarity)",
     horiz = FALSE,
     axes = TRUE,
     cex = 0.7,
     leaflab = "none")  # Don't show default labels

# Add custom colored labels at the bottom
labels <- labels(dend)
pdm_label_colors <- pdm_colors_mapped[order.dendrogram(dend)]
text(1:length(labels), 
     par("usr")[3] - 0.1,  # Position below axis
     labels = labels, 
     col = pdm_label_colors,
     srt = 90,  # Rotate text 90 degrees
     adj = 1,   # Align to right
     cex = 0.6, # Text size
     xpd = TRUE)  # Allow plotting outside figure region

# Add a legend for PDM types
legend("topright", 
       legend = names(pdm_colors),
       fill = pdm_colors,
       title = "Project Delivery Method",
       cex = 0.8,
       bg = "white",
       xpd = TRUE)

# Cut the tree to get 3 clusters (adjust based on your analysis)
k <- 3  # Number of clusters
rect.hclust(hc_ward, k = k, border = 2:(k+1))

# Add cluster labels
clusters <- cutree(hc_ward, k = k)
cluster_centers <- tapply(1:length(clusters), clusters, mean)
text(cluster_centers, par("usr")[4] - 5, 
     labels = paste("Cluster", 1:k), 
     cex = 1.2, 
     font = 2,
     col = 2:(k+1),
     xpd = TRUE)

dev.off()

# Additionally create a dendrogram with branches colored by clusters
dend_colored <- color_branches(dend, k = k)

# Create a second plot with branches colored by cluster
pdf("BOAT2_Dendrogram_Clusters.pdf", width = 12, height = 8)
par(mar = c(8, 4, 4, 10))

plot(dend_colored,
     main = "Hierarchical Clustering of Building Owners by Decision-Making Profiles",
     sub = "Branches Colored by Cluster Assignment",
     xlab = "",
     ylab = "Height (Dissimilarity)",
     horiz = FALSE,
     axes = TRUE,
     leaflab = "none")

# Add cluster rectangles
rect.hclust(hc_ward, k = k, border = 2:(k+1))

# Add cluster labels
text(cluster_centers, par("usr")[4] - 5, 
     labels = paste("Cluster", 1:k), 
     cex = 1.2, 
     font = 2,
     col = 2:(k+1),
     xpd = TRUE)

# Optional: Add PDM labels at the bottom
text(1:length(labels), 
     par("usr")[3] - 0.1,
     labels = labels, 
     col = pdm_label_colors,
     srt = 90,
     adj = 1,
     cex = 0.6,
     xpd = TRUE)

dev.off()

# Output cluster assignments with PDM types
cluster_pdm <- data.frame(
  ID = ids[order.dendrogram(dend)],
  PDM_Type = pdm_types[order.dendrogram(dend)],
  Cluster = clusters[order.dendrogram(dend)]
)

# Save the cluster assignments
write.csv(cluster_pdm, "BOAT2_Cluster_Assignments.csv", row.names = FALSE)

# Create a contingency table of PDM types by cluster
pdm_cluster_table <- table(pdm_types, clusters)
print(pdm_cluster_table)

# Calculate percentages within each cluster
pdm_cluster_percent <- prop.table(pdm_cluster_table, margin = 2) * 100
print(round(pdm_cluster_percent, 1))

# Save the contingency table
write.csv(pdm_cluster_table, "BOAT2_PDM_Cluster_Table.csv")
```

## K-Prototypes Clustering Analysis
This section implements K-prototypes clustering to handle mixed data types (numerical and categorical variables) and compares different clustering solutions.

```{r}
# Remove the outlier with 3700 employees
boat_data_filtered <- boat_data %>%
  filter(Org_Structure_Employees != 3700)

# Define numerical and ordinal variables - using original variables instead of combined ones
numerical_vars <- c(
  "Org_Structure_Employees",
  "Org_Structure_Locations",
  "Org_Structure_Depts",
  "Org_Structure_Layers"
)

ordinal_vars <- c(
  # Use original separate variables instead of combined ones
  "Distribution_Centralization1", 
  "Distribution_Centralization2",
  "Distribution_Formalization1", 
  "Distribution_Formalization2",
  "Style_Technocracy", 
  "Style_Participation1", 
  "Style_Participation2", 
  "Style_Organicity1", 
  "Style_Organicity2", 
  "Style_Coercion1", 
  "Style_Coercion2",
  "Culture_Command", 
  "Culture_Symbolic", 
  "Culture_Rationale", 
  "Culture_Generative", 
  "Culture_Transactive",
  "Flexibility_openness", 
  "Flexibility_Recursiveness",
  "Risk",
  "Environment_Growth", 
  "Environment_Hostile", 
  "Environment_Stable"
)

# Prepare data for clustering and handle missing values
cluster_data <- boat_data_filtered %>%
  select(all_of(c(numerical_vars, ordinal_vars)))

# Print missing value summary
print("Missing values summary:")
print(colSums(is.na(cluster_data)))

# Handle missing values
cluster_data <- cluster_data %>%
  mutate(across(everything(), ~ifelse(is.na(.), median(., na.rm = TRUE), .)))

# Scale all variables
cluster_data_scaled <- scale(cluster_data)

# Determine optimal number of clusters using silhouette method
max_k <- 8
sil_width <- numeric(max_k - 1)

for(k in 2:max_k) {
  pam_fit <- pam(cluster_data_scaled, k = k)
  sil_width[k-1] <- pam_fit$silinfo$avg.width
}

# Plot silhouette widths
silhouette_plot <- ggplot(data.frame(k = 2:max_k, sil_width = sil_width), 
                         aes(x = k, y = sil_width)) +
  geom_line() +
  geom_point() +
  labs(title = "Silhouette Analysis for Optimal k",
       x = "Number of Clusters (k)",
       y = "Average Silhouette Width") +
  theme_minimal()

# Perform PAM clustering with k=2
k <- 2
pam_result <- pam(cluster_data_scaled, k = k)

# Add cluster assignments
cluster_data_scaled <- as.data.frame(cluster_data_scaled)
cluster_data_scaled$cluster <- pam_result$clustering

# Create visualizations

# 1. Numerical variables boxplots by cluster
numerical_long <- cluster_data_scaled %>%
  select(all_of(numerical_vars), cluster) %>%
  pivot_longer(cols = all_of(numerical_vars),
               names_to = "variable",
               values_to = "value")

num_boxplot <- ggplot(numerical_long, 
                     aes(x = factor(cluster), y = value, fill = factor(cluster))) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free_y") +
  labs(title = "Distribution of Numerical Variables by Cluster",
       x = "Cluster",
       y = "Standardized Value") +
  theme_minimal()

# 2. Variables heatmap
cluster_means <- cluster_data_scaled %>%
  group_by(cluster) %>%
  summarise(across(all_of(c(numerical_vars, ordinal_vars)), mean)) %>%
  pivot_longer(-cluster, 
               names_to = "variable", 
               values_to = "value")

heatmap_plot <- ggplot(cluster_means, 
                      aes(x = variable, y = factor(cluster), fill = value)) +
  geom_tile() +
  scale_fill_viridis() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Cluster Characteristics Heatmap",
       x = "Variables",
       y = "Cluster")

# Save results with original (unscaled) values
results_with_originals <- cbind(
  cluster_data,
  cluster = pam_result$clustering
)

write.csv(results_with_originals, "BOAT2_Clustering_Results.csv", row.names = FALSE)

# Save plots
pdf("BOAT2_Clustering_Visualizations.pdf", width = 15, height = 10)
print(silhouette_plot)
print(num_boxplot)
print(heatmap_plot)
dev.off()

# Print summary statistics
print("Cluster Sizes:")
print(table(pam_result$clustering))

# Calculate and print cluster profiles
cluster_profiles <- cluster_data_scaled %>%
  group_by(cluster) %>%
  summarise(across(everything(), 
                  list(mean = ~mean(., na.rm = TRUE),
                       sd = ~sd(., na.rm = TRUE))))

print("\nCluster Profiles:")
print(cluster_profiles)

# Add silhouette information
print("\nSilhouette Information:")
print(summary(pam_result$silinfo$silh[,3]))
```

## Detailed Cluster Analysis and Visualization
This section provides detailed analysis of the clusters and creates various visualizations to understand the characteristics of each cluster.

```{r}
# Create dataset for kprototypes
cluster_data_mixed <- cluster_data
cluster_data_mixed[ordinal_vars] <- lapply(cluster_data_mixed[ordinal_vars], 
                                         function(x) ordered(round(x), levels = 1:5))

# Set random seed for reproducibility
set.seed(123)

# Perform clustering with k=2 and k=3
k2_result <- kproto(cluster_data_mixed, k = 2)
k3_result <- kproto(cluster_data_mixed, k = 3)

# Create comparison dataset
comparison_data <- cluster_data %>%
  mutate(
    cluster_k2 = factor(k2_result$cluster),
    cluster_k3 = factor(k3_result$cluster)
  )

# Create visualizations for numerical variables
numerical_plots <- list()
for(var in numerical_vars) {
  # K=2 boxplot
  p1 <- ggplot(comparison_data, aes_string(x = "cluster_k2", y = var, fill = "cluster_k2")) +
    geom_boxplot() +
    scale_fill_viridis_d() +
    labs(title = paste(var, "- K=2"), x = "Cluster", y = var) +
    theme_minimal()
  
  # K=3 boxplot
  p2 <- ggplot(comparison_data, aes_string(x = "cluster_k3", y = var, fill = "cluster_k3")) +
    geom_boxplot() +
    scale_fill_viridis_d() +
    labs(title = paste(var, "- K=3"), x = "Cluster", y = var) +
    theme_minimal()
  
  numerical_plots[[var]] <- grid.arrange(p1, p2, ncol = 2)
}

# Create cluster characteristics heatmap
cluster_characteristics <- bind_rows(
  # K=2 characteristics
  comparison_data %>%
    group_by(cluster_k2) %>%
    summarise(across(all_of(ordinal_vars), mean)) %>%
    pivot_longer(-cluster_k2, names_to = "variable", values_to = "value") %>%
    mutate(k = "K=2", cluster = as.character(cluster_k2)) %>%
    select(-cluster_k2),
  
  # K=3 characteristics
  comparison_data %>%
    group_by(cluster_k3) %>%
    summarise(across(all_of(ordinal_vars), mean)) %>%
    pivot_longer(-cluster_k3, names_to = "variable", values_to = "value") %>%
    mutate(k = "K=3", cluster = as.character(cluster_k3)) %>%
    select(-cluster_k3)
)

# Create heatmap
characteristics_heatmap <- ggplot(cluster_characteristics, 
                                aes(x = variable, y = paste(k, "Cluster", cluster), 
                                    fill = value)) +
  geom_tile() +
  scale_fill_viridis() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Cluster Characteristics Heatmap",
       x = "Variables",
       y = "Clusters")

# Save visualizations
pdf("Cluster_Analysis_Visualizations.pdf", width = 15, height = 12)

# Print numerical variable plots
for(plot in numerical_plots) {
  grid.arrange(plot)
}

# Print heatmap
print(characteristics_heatmap)

# Print silhouette plots
par(mfrow = c(1,2))
plot(silhouette(k2_result$cluster, daisy(cluster_data_mixed, metric = "gower")), 
     main = "Silhouette Plot - K=2")
plot(silhouette(k3_result$cluster, daisy(cluster_data_mixed, metric = "gower")), 
     main = "Silhouette Plot - K=3")

dev.off()
```

## Dimension Reduction Analysis
This section performs PCA and UMAP analysis to visualize the clustering results in lower dimensions.

```{r}
# Prepare data for PCA - using all individual variables
pca_data <- comparison_data %>%
  select(all_of(c(numerical_vars, ordinal_vars)))

# Perform PCA
pca_result <- prcomp(pca_data, scale. = TRUE)

# Create PCA coordinates with cluster information
pca_coords <- as.data.frame(pca_result$x) %>%
  bind_cols(
    cluster_k2 = factor(comparison_data$cluster_k2),
    cluster_k3 = factor(comparison_data$cluster_k3)
  )

# Create PCA plots
pca_plot_k2 <- ggplot(pca_coords, aes(x = PC1, y = PC2, color = cluster_k2)) +
  geom_point(size = 3, alpha = 0.6) +
  scale_color_viridis_d() +
  labs(title = "PCA Plot - K=2",
       color = "Cluster") +
  theme_minimal() +
  stat_ellipse(level = 0.95)

pca_plot_k3 <- ggplot(pca_coords, aes(x = PC1, y = PC2, color = cluster_k3)) +
  geom_point(size = 3, alpha = 0.6) +
  scale_color_viridis_d() +
  labs(title = "PCA Plot - K=3",
       color = "Cluster") +
  theme_minimal() +
  stat_ellipse(level = 0.95)

# Create UMAP visualization
umap_result <- umap(pca_data)
umap_coords <- as.data.frame(umap_result$layout)
colnames(umap_coords) <- c("UMAP1", "UMAP2")
umap_coords$cluster_k2 <- factor(comparison_data$cluster_k2)
umap_coords$cluster_k3 <- factor(comparison_data$cluster_k3)

# Create UMAP plots
umap_plot_k2 <- ggplot(umap_coords, aes(x = UMAP1, y = UMAP2, color = cluster_k2)) +
  geom_point(size = 3, alpha = 0.6) +
  scale_color_viridis_d() +
  labs(title = "UMAP Plot - K=2",
       color = "Cluster") +
  theme_minimal()

umap_plot_k3 <- ggplot(umap_coords, aes(x = UMAP1, y = UMAP2, color = cluster_k3)) +
  geom_point(size = 3, alpha = 0.6) +
  scale_color_viridis_d() +
  labs(title = "UMAP Plot - K=3",
       color = "Cluster") +
  theme_minimal()

# Create scree plot
variance_explained <- data.frame(
  PC = 1:length(pca_result$sdev),
  Variance = pca_result$sdev^2 / sum(pca_result$sdev^2)
)

scree_plot <- ggplot(variance_explained, aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity") +
  geom_line(aes(y = cumsum(Variance)), color = "red") +
  geom_point(aes(y = cumsum(Variance)), color = "red") +
  scale_y_continuous(labels = percent_format(),
                    sec.axis = sec_axis(~., labels = percent_format())) +
  labs(title = "Scree Plot with Cumulative Variance",
       x = "Principal Component",
       y = "Proportion of Variance Explained") +
  theme_minimal()

# Save dimension reduction visualizations
pdf("Dimension_Reduction_Analysis.pdf", width = 15, height = 12)

# Print PCA plots
grid.arrange(pca_plot_k2, pca_plot_k3, ncol = 2)

# Print UMAP plots
grid.arrange(umap_plot_k2, umap_plot_k3, ncol = 2)

# Print scree plot
print(scree_plot)

dev.off()

# Print PCA summary
print("PCA Summary:")
print(summary(pca_result))

# Print top variable loadings
print("\nTop variable loadings for first 3 PCs:")
loadings <- pca_result$rotation[,1:3]
loadings_df <- as.data.frame(loadings)
loadings_df$Variable <- rownames(loadings_df)
print(loadings_df %>% arrange(desc(abs(PC1))))
```
